---
output:
  md_document:
    variant: markdown_github
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#",
  fig.path = "tools/README-",
  dpi = 250
)
```



# __tropical__ – Tropical Geometry in R

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/tropical)](https://cran.r-project.org/package=tropical)




## Overview

__tropical__ is a new R package that provides users with methods and data structures for [tropical geometry in R](https://en.wikipedia.org/wiki/Tropical_geometry).  Inside R, you load the package with 
```{r load}
library(tropical)
```
See the bottom of this page for how to install __tropical__.


## Basic usage

In tropical geometry, the ordinary operations of addition and multiplication are changed: the sum of two numbers is the minimum of those numbers, and the product of two numbers is their (classical) sum.  To reflect this change, the ordinary plus sign $+$, representing classical addition, is changed to $\oplus$, and ordinary multiplication symbols, $\times$ or $\cdot$, are changed to $\odot$.  As an example, $1 \oplus 5 = 1$, and $1 \odot 5 = 6$.  Together, the set $(\mathbb{R} \cup \infty, \oplus, \odot)$ forms the [algebraic structure](https://en.wikipedia.org/wiki/Algebraic_structure) of a semiring satisfying all ring, and even field, definitions except having additive inverses.

__tropical__ implements $\oplus$ as `%+%` and $\odot$ as `%.%`.  For example:
```{r plus-and-times}
1 %+% 5
1 %.% 5
```
As in base R, these operations are vectorized:
```{r plus-and-times-vectorized}
1:3 %+% 3:1
1:3 %.% 3:1
```
Exponentiation is also defined as iterated multiplication ($\oplus$), but only when natural number exponents are given, so that $5^{3}$ (sometimes written $5^{\odot 3}$) is $5^{3} = 5 \odot 5 \odot 5 = 5 + 5 + 5 = 3 \times 5 = 15$.  Example:
```{r exponentiation}
5 %^% 3
```
This operation is vectorized, too.

Matrix multiplication is also defined through inner products of vectors.  For two vectors $\textbf{x}$ and $\textbf{y}$ of the same length, their inner product is $\textbf{x}'\textbf{y} = \bigoplus (x_{k} \odot y_{k}) = \min\{x_{k}+y_{k}\}$.  In __tropical__, this is given by `%..%` (which is the tropical analogue of R's ordinary matrix multiplication operator `%*%`):
```{r inner-product}
1:3 %..% 4:6
```
This is $\min \{1+4, 2+5, 3+6 \} = 5$.  From there, matrix multiplication is defined in the standard way, so that for two matrices $\textbf{A}$ and $\textbf{B}$, their product $\textbf{C} = \textbf{AB}$ has elements $c_{i,j} = \textbf{a}_{i}'\textbf{b}_{j}$, where $\textbf{a}_{i}$ is the $i$th row of $\textbf{A}$ and $\textbf{b}_{j}$ is the $j$th column of $\textbf{B}$. Example:
```{r matrix-mult}
(m1 <- matrix(1:6, 2, 3))
(m2 <- matrix(6:1, 3, 2))
m1 %..% m2
```

Sometimes tropical arithmetic is defined with $\oplus$ interpreted as the maximum of the two numbers, not the minimum.  To change __tropical__'s default setting, just type:
```{r set-plus-max}
set_plus_max()
1:3 %+% 3:1
```
Once you use this function, all operations will have the appropriate corresponding changes. You can change it back with 
```{r set-plus-min}
set_plus_min()
1:3 %+% 3:1
```


## Acknowledgements

This material is based upon work supported by the National Science Foundation under Grant Nos. [1321794](https://nsf.gov/awardsearch/showAward?AWD_ID=1321794) and [1622449](https://nsf.gov/awardsearch/showAward?AWD_ID=1622449).

## Installation

Here's how you can install the current _developmental_ version of __tropical__:

```R
if(!requireNamespace("devtools")) install.packages("devtools")
devtools::install_github("dkahle/tropical")
```
